---
title: "Title"
author: "Author"
date: '`r Sys.Date()`'
output:
  md_document: 
    variant: markdown_github
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=F, warning=FALSE, message=FALSE)
```


```{r}
setwd("~/git/subgraph/mgc_based/")

require('igraph')
require('ggplot2')
require('fields')

options(max.print=100)

listGs<- list.files(path = "../graphml/", pattern = "*.graphml")

#read in covariates and graph list
#find those with common ids, sort by id

covariates<- read.csv("../covariates/predictors.csv",stringsAsFactors = F)
ids <- unlist( lapply(listGs,function(x)strtrim(x,6)))
common_id<- intersect(covariates$RUNNO , ids)

covariates <- covariates[covariates$RUNNO%in%common_id,]
covariates <- covariates[order(covariates$RUNNO),]  

listGs<- listGs[ids%in%common_id]
listGs<- listGs[order(listGs)]

graphList<- lapply(listGs, function(x){
  read.graph( file = paste("../graphml/",x,sep = ""),format = "graphml")
})

AdjacencyList<- lapply(graphList, function(x){
  get.adjacency(x)
})

n = nrow(AdjacencyList[[1]])
half = c(1:(n/2))
reordered = c( half*2-1,half*2)


AdjacencyList<- lapply(AdjacencyList, function(x){
  x[reordered,reordered]
})

HemisphereList<- lapply(graphList, function(x){
  get.vertex.attribute(x,name="hemisphere")
})

DegreeList<- lapply(AdjacencyList, function(x){
  rowSums(as.matrix(x))
  })

####

AdjacencyListPick =  list()
for(i in 1:length(AdjacencyList)){
  if(covariates$GENOTYPE[i]>=1){
    AdjacencyListPick[[length(AdjacencyListPick)+1]] = AdjacencyList[[i]]
  }
}

GENOTYPEPick = covariates$GENOTYPE[covariates$GENOTYPE>=1]

m = length(GENOTYPEPick)
```


#Raw Adjacency Plots

Raw Adjacency Plots for Genotype 1

```{r}


#genotype 1
geno1 = c(1:m)[GENOTYPEPick == 1]
par(mfrow=c(1,2))
for(i in geno1){
    img = 1-as.matrix(AdjacencyListPick[[i]])
    image(img, col  = gray((0:32)/32),main=i)
}

```


Raw Adjacency Plots for Genotype 2

```{r}
#genotype 2
geno2 = c(1:m)[GENOTYPEPick == 2]
par(mfrow=c(1,2))
for(i in geno2){
    img = 1-as.matrix(AdjacencyListPick[[i]])
    image(img, col  = gray((0:32)/32),main=i)
}


```


#Average Adjacency Plots

Genotype 1

```{r}

Asum = matrix(0, n,n)
for(i in geno1){
    Asum = Asum + as.matrix(AdjacencyListPick[[i]])
}

avgGeno1 = Asum/length(geno1)

image.plot(avgGeno1)

```
Genotype 2

```{r}



Asum = matrix(0, n,n)
for(i in geno2){
    Asum = Asum + as.matrix(AdjacencyListPick[[i]])
}

avgGeno2 = Asum/length(geno2)

image.plot(avgGeno2)


```


#Degree Distribution

Plotted over two genotypes

```{r}

DegreeList<- lapply(AdjacencyList, function(x){
  rowSums(as.matrix(x))
  })

DegreeListMat = do.call("cbind", DegreeList)

DegreeListPick = DegreeListMat[,covariates$GENOTYPE>=1]



df = data.frame("degree"= c(DegreeListPick), "id"= as.factor(rep(c(1:m),each=n)),"genotype"= as.factor(rep(GENOTYPEPick,each=n)))

ggplot(data=df, aes(degree,fill=id))+ geom_histogram(bins = 30,position = "identity",alpha=0.5)+ theme_bw()+facet_grid(~genotype)

```


#Betweenness Centrality Distribution

Plotted over two genotypes

```{r}


DegreeList<- lapply(graphList, function(x){
  betweenness(x)
})

DegreeListMat = do.call("cbind", DegreeList)

DegreeListPick = DegreeListMat[,covariates$GENOTYPE>=1]

df = data.frame("betweenness"= c(DegreeListPick), "id"= as.factor(rep(c(1:m),each=n)),"genotype"= as.factor(rep(GENOTYPEPick,each=n)))

ggplot(data=df, aes(betweenness,fill=id))+ geom_histogram(bins = 30,position = "identity",alpha=0.5)+ theme_bw()+facet_grid(~genotype)

```




#Left and Right

```{r}

left_idx = c(1: (n/2))
right_idx = c((n/2+1):n)

LRSumList = lapply( AdjacencyListPick, function(x){
  A = matrix(0,2,2)
  A[1,1] = sum(x[left_idx,left_idx])
  A[1,2] = sum(x[left_idx,right_idx])
  A[2,2] = sum(x[right_idx,right_idx])
  A[2,1] = A[1,2]
  A
})


```


Genotype1

```{r}
par(mfrow=c(1,2))
for(i in geno1){
    img = as.matrix(LRSumList[[i]])
    image.plot(img,main=i, zlim=c(0,7000))
}

```


Genotype2

```{r}
par(mfrow=c(1,2))
for(i in geno2){
    img = as.matrix(LRSumList[[i]])
    image.plot(img,main=i, zlim=c(0,7000))
}

```



Average

Genotype 1

```{r}

Asum = matrix(0, 2,2)
for(i in geno1){
    Asum = Asum + as.matrix(LRSumList[[i]])
}

avgGeno1 = Asum/length(geno1)

image.plot(avgGeno1, zlim=c(0,7000))

```
Genotype 2

```{r}



Asum = matrix(0, 2,2)
for(i in geno2){
    Asum = Asum + as.matrix(LRSumList[[i]])
}

avgGeno2 = Asum/length(geno2)

image.plot(avgGeno2, zlim=c(0,7000))

```


Plot the genotype1(black) vs genotype2(red) in the edge count in each block.

```{r}

plot(c(avgGeno1),ylim=c(0,6000))
lines(c(avgGeno2),col="red",type="p")

```