---
title: "Title"
author: "Author"
date: '`r Sys.Date()`'
output:
  md_document: 
    variant: markdown_github
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='FigsEDA/',
                      echo=F, warning=FALSE, message=FALSE)
```


```{r}
setwd("~/git/subgraph/mgc_based/")

require('igraph')
require('ggplot2')
require('fields')

options(max.print=100)

listGs<- list.files(path = "../graphml/", pattern = "*.graphml")

#read in covariates and graph list
#find those with common ids, sort by id

covariates<- read.csv("../covariates/predictors.csv",stringsAsFactors = F)
ids <- unlist( lapply(listGs,function(x)strtrim(x,6)))
common_id<- intersect(covariates$RUNNO , ids)

covariates <- covariates[covariates$RUNNO%in%common_id,]
covariates <- covariates[order(covariates$RUNNO),]  

listGs<- listGs[ids%in%common_id]
listGs<- listGs[order(listGs)]

graphList<- lapply(listGs, function(x){
  read.graph( file = paste("../graphml/",x,sep = ""),format = "graphml")
})

AdjacencyList<- lapply(graphList, function(x){
  get.adjacency(x)
})

n = nrow(AdjacencyList[[1]])
half = c(1:(n/2))
reordered = c( half*2-1,half*2)


AdjacencyList<- lapply(AdjacencyList, function(x){
  x[reordered,reordered]
})

HemisphereList<- lapply(graphList, function(x){
  get.vertex.attribute(x,name="hemisphere")
})

DegreeList<- lapply(AdjacencyList, function(x){
  rowSums(as.matrix(x))
  })

####

AdjacencyListPick =  list()
for(i in 1:length(AdjacencyList)){
  if(covariates$GENOTYPE[i]>=1){
    AdjacencyListPick[[length(AdjacencyListPick)+1]] = AdjacencyList[[i]]
  }
}

GENOTYPEPick = covariates$GENOTYPE[covariates$GENOTYPE>=1]

SEXPick = covariates$GENDER[covariates$GENOTYPE>=1]

m = length(GENOTYPEPick)
```


#Raw Adjacency Plots

Raw Adjacency Plots for Genotype 1

```{r}


#genotype 1
geno1 = c(1:m)[GENOTYPEPick == 1]
par(mfrow=c(1,2))
for(i in geno1){
    img = 1-as.matrix(AdjacencyListPick[[i]])
    image(img, col  = gray((0:32)/32),main=i)
}

```


Raw Adjacency Plots for Genotype 2

```{r}
#genotype 2
geno2 = c(1:m)[GENOTYPEPick == 2]
par(mfrow=c(1,2))
for(i in geno2){
    img = 1-as.matrix(AdjacencyListPick[[i]])
    image(img, col  = gray((0:32)/32),main=i)
}


```


#Average Adjacency Plots

Genotype 1

```{r}

Asum = matrix(0, n,n)
for(i in geno1){
    Asum = Asum + as.matrix(AdjacencyListPick[[i]])
}

avgGeno1 = Asum/length(geno1)

image.plot(avgGeno1)

```
Genotype 2

```{r}



Asum = matrix(0, n,n)
for(i in geno2){
    Asum = Asum + as.matrix(AdjacencyListPick[[i]])
}

avgGeno2 = Asum/length(geno2)

image.plot(avgGeno2)


```


#Degree Distribution

Plotted over two genotypes

```{r}

DegreeList<- lapply(AdjacencyList, function(x){
  rowSums(as.matrix(x))
  })

DegreeListMat = do.call("cbind", DegreeList)

DegreeListPick = DegreeListMat[,covariates$GENOTYPE>=1]



df = data.frame("degree"= c(DegreeListPick), "id"= as.factor(rep(c(1:m),each=n)),"genotype"= as.factor(rep(GENOTYPEPick,each=n)),"sex"=as.factor(rep(SEXPick,each=n)))

ggplot(data=df, aes(degree,fill=id))+ geom_histogram(bins = 30,position = "identity",alpha=0.5)+ theme_bw()+facet_grid(~genotype)

```

Plotted over two sex
```{r}
ggplot(data=df, aes(degree,fill=id))+ geom_histogram(bins = 30,position = "identity",alpha=0.5)+ theme_bw()+facet_grid(~sex)

```


#Betweenness Centrality Distribution

Plotted over two genotypes

```{r}


DegreeList<- lapply(graphList, function(x){
  betweenness(x)
})

DegreeListMat = do.call("cbind", DegreeList)

DegreeListPick = DegreeListMat[,covariates$GENOTYPE>=1]

df = data.frame("betweenness"= c(DegreeListPick), "id"= as.factor(rep(c(1:m),each=n)),"genotype"= as.factor(rep(GENOTYPEPick,each=n)),"sex"=as.factor(rep(SEXPick,each=n)))

ggplot(data=df, aes(betweenness,fill=id))+ geom_histogram(bins = 30,position = "identity",alpha=0.5)+ theme_bw()+facet_grid(~genotype)

```


Plotted over two sex

```{r}
ggplot(data=df, aes(betweenness,fill=id))+ geom_histogram(bins = 30,position = "identity",alpha=0.5)+ theme_bw()+facet_grid(~sex)

```


```{r}

left_idx = c(1: (n/2))
right_idx = c((n/2+1):n)

LRSumList = lapply( AdjacencyListPick, function(x){
  A = matrix(0,2,2)
  A[1,1] = sum(x[left_idx,left_idx])
  A[1,2] = sum(x[left_idx,right_idx])
  A[2,2] = sum(x[right_idx,right_idx])
  A[2,1] = A[1,2]
  A
})


```


<!-- Genotype1 -->

<!-- ```{r} -->
<!-- par(mfrow=c(1,2)) -->
<!-- for(i in geno1){ -->
<!--     img = as.matrix(LRSumList[[i]]) -->
<!--     image.plot(img,main=i, zlim=c(0,7000)) -->
<!-- } -->

<!-- ``` -->


<!-- Genotype2 -->

<!-- ```{r} -->
<!-- par(mfrow=c(1,2)) -->
<!-- for(i in geno2){ -->
<!--     img = as.matrix(LRSumList[[i]]) -->
<!--     image.plot(img,main=i, zlim=c(0,7000)) -->
<!-- } -->

<!-- ``` -->



<!-- Average -->

<!-- Genotype 1 -->

<!-- ```{r} -->

<!-- Asum = matrix(0, 2,2) -->
<!-- for(i in geno1){ -->
<!--     Asum = Asum + as.matrix(LRSumList[[i]]) -->
<!-- } -->

<!-- avgGeno1 = Asum/length(geno1) -->

<!-- image.plot(avgGeno1, zlim=c(0,7000)) -->

<!-- ``` -->
<!-- Genotype 2 -->

<!-- ```{r} -->



<!-- Asum = matrix(0, 2,2) -->
<!-- for(i in geno2){ -->
<!--     Asum = Asum + as.matrix(LRSumList[[i]]) -->
<!-- } -->

<!-- avgGeno2 = Asum/length(geno2) -->

<!-- image.plot(avgGeno2, zlim=c(0,7000)) -->

<!-- ``` -->



```{r}
z.test2sam = function(a, b, var.a, var.b){
   n.a = length(a)
   n.b = length(b)
   zeta = (mean(a) - mean(b)) / (sqrt(var.a/n.a + var.b/n.b))
   return(zeta)
}

```

# total edges
```{r}

total_edges<-sapply(LRSumList, function(x){
  sum(x)
})


df = data.frame("total_edges"= c(total_edges), "id"= as.factor((c(1:m))),"genotype"= as.factor(GENOTYPEPick),"sex"= as.factor(SEXPick))


```


```{r}


n1 = sum(df$genotype==1)
p1 = sum(total_edges[df$genotype==1]) / (n^2) /n1

sigma2.x1 = (n^2/4)*p1*(1-p1)

n2 = sum(df$genotype==2)
p2 = sum(total_edges[df$genotype==2]) / (n^2) /n2

sigma2.x2 = (n^2/4)*p2*(1-p2)


```



Over genotype 
```{r}

ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = total_edges[df$genotype==1]
x2 = total_edges[df$genotype==2]

wilcox.test(x1,x2, alternative = "greater")
kruskal.test(df$total_edges~as.factor(df$genotype))


```


Over sex 
```{r}

ggplot(data=df, aes(x= 1, y=total_edges,col=sex,group=sex))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~sex)


x1 = total_edges[df$sex==1]
x2 = total_edges[df$sex==2]

wilcox.test(x1,x2, alternative = "two.sided")
kruskal.test(df$total_edges~as.factor(df$sex))

```

# edges in left hemisphere
```{r}

total_edges<-sapply(LRSumList, function(x){
  sum(x[1,1])
})


df = data.frame("total_edges"= c(total_edges), "id"= as.factor((c(1:m))),"genotype"= as.factor((GENOTYPEPick)), "sex"=as.factor(SEXPick))


# t.test(x1,x2,alternative = "greater")




```


Over genotype 
```{r}

ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = total_edges[df$genotype==1]
x2 = total_edges[df$genotype==2]

wilcox.test(x1,x2, alternative = "greater")
kruskal.test(df$total_edges~as.factor(df$genotype))

```


Over sex 
```{r}

ggplot(data=df, aes(x= 1, y=total_edges,col=sex,group=sex))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~sex)


x1 = total_edges[df$sex==1]
x2 = total_edges[df$sex==2]

wilcox.test(x1,x2, alternative = "two.sided")
kruskal.test(df$total_edges~as.factor(df$sex))

```

# edges in right hemisphere
```{r}

total_edges<-sapply(LRSumList, function(x){
  sum(x[2,2])
})


df = data.frame("total_edges"= c(total_edges), "id"= as.factor((c(1:m))),"genotype"= as.factor((GENOTYPEPick)), "sex"=as.factor(SEXPick))


```


```{r}


n1 = sum(df$genotype==1)
p1 = sum(total_edges[df$genotype==1]) / (n^2/4) /n1

sigma2.x1 = (n^2/4)*p1*(1-p1)

n2 = sum(df$genotype==2)
p2 = sum(total_edges[df$genotype==2]) / (n^2/4) /n2

sigma2.x2 = (n^2/4)*p2*(1-p2)



```


Over genotype 
```{r}

ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = total_edges[df$genotype==1]
x2 = total_edges[df$genotype==2]

wilcox.test(x1,x2, alternative = "greater")
kruskal.test(df$total_edges~as.factor(df$genotype))

```


Over sex 
```{r}

ggplot(data=df, aes(x= 1, y=total_edges,col=sex,group=sex))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~sex)


x1 = total_edges[df$sex==1]
x2 = total_edges[df$sex==2]

wilcox.test(x1,x2, alternative = "two.sided")
kruskal.test(df$total_edges~as.factor(df$sex))

```

# edges across left and right 
```{r}

total_edges<-sapply(LRSumList, function(x){
  sum(x[1,2])
})


df = data.frame("total_edges"= c(total_edges), "id"= as.factor((c(1:m))),"genotype"= as.factor((GENOTYPEPick)), "sex"=as.factor(SEXPick))


```

```{r}


n1 = sum(df$genotype==1)
p1 = sum(total_edges[df$genotype==1]) / (n^2/4) /n1

sigma2.x1 = (n^2/4)*p1*(1-p1)

n2 = sum(df$genotype==2)
p2 = sum(total_edges[df$genotype==2]) / (n^2/4) /n2

sigma2.x2 = (n^2/4)*p2*(1-p2)


```



Over genotype 
```{r}

ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = total_edges[df$genotype==1]
x2 = total_edges[df$genotype==2]

wilcox.test(x1,x2, alternative = "greater")

kruskal.test(df$total_edges~as.factor(df$genotype))

```


Over sex 
```{r}

ggplot(data=df, aes(x= 1, y=total_edges,col=sex,group=sex))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~sex)


x1 = total_edges[df$sex==1]
x2 = total_edges[df$sex==2]

wilcox.test(x1,x2, alternative = "two.sided")

kruskal.test(df$total_edges~as.factor(df$sex))

```