---
title: "Data Analysis"
author: "Author"
date: '`r Sys.Date()`'
output:
  md_document: 
    variant: markdown_github
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='FigsReport/',
                      echo=F, warning=FALSE, message=F)
```

```{r}
setwd("~/git/subgraph/mgc_based/")

require('igraph')
require('ggplot2')
require('fields')

options(max.print=100)

listGs<- list.files(path = "../graphml/", pattern = "*.graphml")

#read in covariates and graph list
#find those with common ids, sort by id

covariates<- read.csv("../covariates/predictors.csv",stringsAsFactors = F)
ids <- unlist( lapply(listGs,function(x)strtrim(x,6)))
common_id<- intersect(covariates$RUNNO , ids)

covariates <- covariates[covariates$RUNNO%in%common_id,]
covariates <- covariates[order(covariates$RUNNO),]  

listGs<- listGs[ids%in%common_id]
listGs<- listGs[order(listGs)]

graphList<- lapply(listGs, function(x){
  read.graph( file = paste("../graphml/",x,sep = ""),format = "graphml")
})

AdjacencyList<- lapply(graphList, function(x){
  get.adjacency(x)
})


n = nrow(AdjacencyList[[1]])
half = c(1:(n/2))
reordered = c( half*2-1,half*2)

TriangleList <-  lapply(graphList, function(x){
  a=count_triangles(x)
  a[reordered]
})


CluscoeffList<- lapply(graphList, function(x){
  a=transitivity(x,"local")
  a[reordered]
})

AdjacencyList<- lapply(AdjacencyList, function(x){
  x[reordered,reordered]
})

HemisphereList<- lapply(graphList, function(x){
  get.vertex.attribute(x,name="hemisphere")
})

DegreeList<- lapply(AdjacencyList, function(x){
  rowSums(as.matrix(x))
  })

####

AdjacencyListPick =  list()
TriangleListPick =list()
CluscoeffListPick = list()
for(i in 1:length(AdjacencyList)){
  if(covariates$GENOTYPE[i]>=1){
    AdjacencyListPick[[length(AdjacencyListPick)+1]] = AdjacencyList[[i]]
    TriangleListPick[[length(TriangleListPick)+1]] = TriangleList[[i]]
    CluscoeffListPick[[length(CluscoeffListPick)+1]] = CluscoeffList[[i]]
  }
}


DegreeListPick<- lapply(AdjacencyListPick, function(x){
  rowSums(as.matrix(x))
  })



GENOTYPEPick = covariates$GENOTYPE[covariates$GENOTYPE>=1]

SEXPick = covariates$GENDER[covariates$GENOTYPE>=1]

m = length(GENOTYPEPick)
```


# Difference between two genotypes

We focus on comparing subjects with genotype 1 (n1 = 9) and genotype 2 (n2=12).


```{r}


left_idx = c(1: (n/2))
right_idx = c((n/2+1):n)

LRSumList = lapply( AdjacencyListPick, function(x){
  A = matrix(0,2,2)
  A[1,1] = sum(x[left_idx,left_idx])
  A[1,2] = sum(x[left_idx,right_idx])
  A[2,2] = sum(x[right_idx,right_idx])
  A[2,1] = A[1,2]
  A
})



```

In the total number of edges on the whole brain, there seems to be a difference between two genotypes.

```{r}

total_edges<-sapply(LRSumList, function(x){
  sum(x)
})

df = data.frame("total_edges"= c(total_edges), "id"= as.factor((c(1:m))),"genotype"= as.factor(GENOTYPEPick),"sex"= as.factor(SEXPick))
n1 = sum(df$genotype==1)
n2 = sum(df$genotype==2)

ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = total_edges[df$genotype==1]
x2 = total_edges[df$genotype==2]

wilcox.test(x1,x2, alternative = "greater")
kruskal.test(df$total_edges~as.factor(df$genotype))


```



 *cross-hemisphere* connections show a significant difference.

```{r}
total_edges<-sapply(LRSumList, function(x){
  sum(x[1,2])
})


df = data.frame("total_edges"= c(total_edges), "id"= as.factor((c(1:m))),"genotype"= as.factor((GENOTYPEPick)), "sex"=as.factor(SEXPick))

ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = total_edges[df$genotype==1]
x2 = total_edges[df$genotype==2]

wilcox.test(x1,x2, alternative = "greater")

kruskal.test(df$total_edges~as.factor(df$genotype))

```


 within-L-hemisphere connections show a significant difference.

```{r}
total_edges<-sapply(LRSumList, function(x){
  (x[1,1])
})


df = data.frame("total_edges"= c(total_edges), "id"= as.factor((c(1:m))),"genotype"= as.factor((GENOTYPEPick)), "sex"=as.factor(SEXPick))

ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = total_edges[df$genotype==1]
x2 = total_edges[df$genotype==2]

wilcox.test(x1,x2, alternative = "greater")

kruskal.test(df$total_edges~as.factor(df$genotype))

```


 within-R-hemisphere connections do NOT show a significant difference.

```{r}
total_edges<-sapply(LRSumList, function(x){
  (x[2,2])
})


df = data.frame("total_edges"= c(total_edges), "id"= as.factor((c(1:m))),"genotype"= as.factor((GENOTYPEPick)), "sex"=as.factor(SEXPick))

ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = total_edges[df$genotype==1]
x2 = total_edges[df$genotype==2]

wilcox.test(x1,x2, alternative = "greater")

kruskal.test(df$total_edges~as.factor(df$genotype))

```




As a reference, there does not seem to be a clear difference between two sexes:

```{r}

ggplot(data=df, aes(x= 1, y=total_edges,col=sex,group=sex))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~sex)


x1 = total_edges[df$sex==1]
x2 = total_edges[df$sex==2]

wilcox.test(x1,x2, alternative = "two.sided")
kruskal.test(df$total_edges~as.factor(df$sex))

```




#Jointly embedding the graph in low dimensional space

The graphs are jointly embedded into low-dimensional space (with d=10). This is equivalent to PCA, except the principle components (factors) are shared among all subjects. The following plots the top 10 factors.
```{r}

load(file= "embedding.Rda")

Dlist = sapply(stiefelDecomp$D_list,function(x){
  diag(x)
})

D = t(Dlist)


U = stiefelDecomp$U0[,1:10]

df = data.frame("idx"=c(1:n),"h"=c(U),"dim"=as.factor(rep(c(1:10),each=n)))

ggplot(data=df) + geom_line(aes(x=idx,y=h))+facet_wrap(~dim,scale="free") + 
  theme(axis.title=element_blank(), 
        axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(strip.text.y = element_text(angle = 0, vjust=0.2, hjust=0)) +
  theme(axis.title=element_blank(), 
        axis.text.x = element_blank(), axis.ticks = element_blank()) +
  theme(strip.text.x = element_text(angle = 0, vjust=0.2, hjust=0)) 

```


 Each graph is a transform of the linear combination of the 10 eigenmaps.
 
 
The left shows the raw estimate, the right shows the clustered estimate. The vertices are sorted by the vertice index in each hemisphere.

 
```{r}

par(mfrow=c(1,2))

for(i in 1:10)
{
  image.plot(U[,i]%*%t(U[,i]))
  km6 = kmeans(U[,i],centers = 2)
  U6fitted = km6$centers[km6$cluster]
  U6fitted2 = U6fitted %*%t(U6fitted)
  image.plot(U6fitted2)
}

```



#Omni-embedding the graph in low dimensional space

We use another embedding named omni-embedding. Each graph is embedded into 2-dimensional space. The difference is that the principle components (factors) are NOT shared across subjects.

We found there is some separation of the genotypes on the 2-dimensional representation of one vertex. Vertex 111 and its couterpart in the R hemisphere is shown.

```{r}
load(file ="omni_embedding.Rda")

Xhat = Xhat.out$Xhat
df0 = data.frame( "x1"=Xhat[,1] ,"x2"=Xhat[,2], "id"= as.factor(rep(c(1:m),each=n)),"genotype"= as.factor(rep(GENOTYPEPick,each=n)),"sex"=as.factor(rep(SEXPick,each=n)), "vertex"= (rep(c(1:n),m)))

df1 = df0[df0$vertex %in% c(111, 277),]
ggplot(df1, aes(x=x1, y=x2,col=genotype)) +  geom_point() + facet_wrap(~vertex)
```



In the degree distribution of vertex 53 over 18 subjects, there is some difference between two genotypes:


```{r}

degree111 = sapply(DegreeListPick, function(x)x[53])

df1 = data.frame( "degree" =  degree111, "genotype" = as.factor(GENOTYPEPick), "sex" = SEXPick)

ggplot(data=df1, aes(x= 1, y=degree,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = df1$degree[df1$genotype ==1]
x2 = df1$degree[df1$genotype==2]

wilcox.test(x1,x2, alternative = "greater")
kruskal.test(df1$degree~as.factor(df1$genotype))

```

In the degree distribution of vertex 111 over 18 subjects, there is some difference between two genotypes:


```{r}

degree111 = sapply(DegreeListPick, function(x)x[111])

df1 = data.frame( "degree" =  degree111, "genotype" = as.factor(GENOTYPEPick), "sex" = SEXPick)

ggplot(data=df1, aes(x= 1, y=degree,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = df1$degree[df1$genotype ==1]
x2 = df1$degree[df1$genotype==2]

wilcox.test(x1,x2, alternative = "greater")
kruskal.test(df1$degree~as.factor(df1$genotype))

```



#Other things we assessed but did not find meaningful signal

1. Difference between total connections within either hemisphere, with respect to sex
2. Difference between Total connections across hemipheres, with respect to sex
3. Screening individual vertex without embedding.
