---
title: "Data Analysis"
author: "Author"
date: '`r Sys.Date()`'
output:
  md_document: 
    variant: markdown_github
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='FigROItest/',
                      echo=T, warning=FALSE, message=T)
```

```{r}
setwd("~/git/subgraph/mgc_based/")

require('igraph')
require('ggplot2')
require('fields')

options(max.print=100)

listGs<- list.files(path = "../graphml/", pattern = "*.graphml")

#read in covariates and graph list
#find those with common ids, sort by id

covariates<- read.csv("../covariates/predictors.csv",stringsAsFactors = F)
ids <- unlist( lapply(listGs,function(x)strtrim(x,6)))
common_id<- intersect(covariates$RUNNO , ids)

covariates <- covariates[covariates$RUNNO%in%common_id,]
covariates <- covariates[order(covariates$RUNNO),]  

listGs<- listGs[ids%in%common_id]
listGs<- listGs[order(listGs)]

graphList<- lapply(listGs, function(x){
  read.graph( file = paste("../graphml/",x,sep = ""),format = "graphml")
})

AdjacencyList<- lapply(graphList, function(x){
  get.adjacency(x)
})


n = nrow(AdjacencyList[[1]])
half = c(1:(n/2))
reordered = c( half*2-1,half*2)

TriangleList <-  lapply(graphList, function(x){
  a=count_triangles(x)
  a[reordered]
})


CluscoeffList<- lapply(graphList, function(x){
  a=transitivity(x,"local")
  a[reordered]
})

AdjacencyList<- lapply(AdjacencyList, function(x){
  x[reordered,reordered]
})

HemisphereList<- lapply(graphList, function(x){
  get.vertex.attribute(x,name="hemisphere")
})

DegreeList<- lapply(AdjacencyList, function(x){
  rowSums(as.matrix(x))
  })

####

AdjacencyListPick =  list()
TriangleListPick =list()
CluscoeffListPick = list()
for(i in 1:length(AdjacencyList)){
  if(covariates$GENOTYPE[i]>=1){
    AdjacencyListPick[[length(AdjacencyListPick)+1]] = AdjacencyList[[i]]
    TriangleListPick[[length(TriangleListPick)+1]] = TriangleList[[i]]
    CluscoeffListPick[[length(CluscoeffListPick)+1]] = CluscoeffList[[i]]
  }
}


DegreeListPick<- lapply(AdjacencyListPick, function(x){
  rowSums(as.matrix(x))
  })



GENOTYPEPick = covariates$GENOTYPE[covariates$GENOTYPE>=1]

SEXPick = covariates$GENDER[covariates$GENOTYPE>=1]

m = length(GENOTYPEPick)
```

```{r}

pick1= c(63, 52, 60,62,164,166,121,123,74,158,92,93,94,95,96)

sg = lapply(AdjacencyListPick, function(x){
  pick = c(pick1, pick1+166)
  as.matrix(x[pick,pick])
})

image.plot(sg[[4]])

```


```{r}

plot_and_test <- function(total_edges){
  
  df = data.frame("total_edges"=total_edges, "id"= as.factor((c(1:m))),"genotype"= as.factor(GENOTYPEPick),"sex"= as.factor(SEXPick))

  p = ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+ geom_jitter(width = 0.5, height = 0)+geom_boxplot(alpha=0.5)+facet_grid(~genotype)
  print(p)
  
  x1 = df$total_edges[df$genotype==1]
  x2 = df$total_edges[df$genotype==2]
  
  print(wilcox.test(x1,x2, alternative = "two.sided"))
  print(kruskal.test(df$total_edges~as.factor(df$genotype)))

}

check4<- function(pick){
  
  #total
  pick_L= pick
  pick_R= pick+n_sg/2
  pick_F = c(pick_L,pick_R)
  
  print("Both hemispheres")
  #whole
  total_edges = sapply(sg, function(x) sum(x[pick_F,pick_F]))
  plot_and_test(total_edges)
    print("Left hemispheres")

  #Left
  total_edges = sapply(sg, function(x) sum(x[pick_L,pick_L]))
  plot_and_test(total_edges)
    print("Right hemispheres")

  #right
  total_edges = sapply(sg, function(x) sum(x[pick_R,pick_R]))
  plot_and_test(total_edges)
    print("Between hemispheres")

  #across
  total_edges = sapply(sg, function(x) sum(x[pick_L,pick_R]))
  plot_and_test(total_edges)


}


```
List of regions:

1.
septum 63
hippocampus 52
hypothalamus circuit  60 62 164 166

2.
fimbria-fornix 121 123

3.
possibly thalamic nuclei  74 158

4.
cerebellum 92 93 94 95 96

#Total connectivities in all 4 regions 

```{r}
n_sg= nrow(sg[[1]])
pick2 = 1:(n_sg/2)
check4(pick2)
```


##region 1:
septum 63
hippocampus 52
hypothalamus circuit  60 62 164 166


### septum and hippocampus
```{r}
pick2 = c(1:2)
check4(pick2)
```


### hypothalamus circuit
```{r}
pick2 = c(3:6)
check4(pick2)
```



### septum and hippocampus + hypothalamus circuit
```{r}
pick2 = c(1:6)
check4(pick2)
```



total_edges = sapply(sg, function(x) sum(x[pick2, (pick2+n_sg/2) ]))

df = data.frame("total_edges"=total_edges, "id"= as.factor((c(1:m))),"genotype"= as.factor(GENOTYPEPick),"sex"= as.factor(SEXPick))

ggplot(data=df, aes(x= 1, y=total_edges,col=genotype,group=genotype))+ geom_point()+geom_jitter()+geom_boxplot(alpha=0.5)+facet_grid(~genotype)


x1 = df$total_edges[df$genotype==1]
x2 = df$total_edges[df$genotype==2]

wilcox.test(x1,x2, alternative = "two.sided")
kruskal.test(df$total_edges~as.factor(df$genotype))


## region 2: fimbria-fornix 121 123
```{r}
pick2 = c(7:8)
check4(pick2)
```


##region 3: possibly thalamic nuclei  74 158
```{r}
pick2 = c(9:10)
check4(pick2)
```


##region 3: cerebellum

```{r}
pick2 = c(11:15)
check4(pick2)
```
