---
output:
  pdf_document: default
  html_document: default
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
require('igraph')
require('ggplot2')
require('reshape')
require('lsr')
require('transport')

setwd("c:/Users/leo/git/ABadea/")

listGs<- list.files(path = "./graphml/", pattern = "*.graphml")

#read in covariates and graph list
#find those with common ids, sort by id

covariates<- read.csv("./covariates/predictors.csv",stringsAsFactors = F)
ids <- unlist( lapply(listGs,function(x)strtrim(x,6)))
common_id<- intersect(covariates$RUNNO , ids)

covariates <- covariates[covariates$RUNNO%in%common_id,]
covariates <- covariates[order(covariates$RUNNO),]  

listGs<- listGs[ids%in%common_id]
listGs<- listGs[order(listGs)]

graphList<- lapply(listGs, function(x){
  read.graph( file = paste("./graphml/",x,sep = ""),format = "graphml")
})

AdjacencyList<- lapply(graphList, function(x){
  get.adjacency(x)
})

HemisphereList<- lapply(graphList, function(x){
  get.vertex.attribute(x,name="hemisphere")
})

DegreeList<- lapply(AdjacencyList, function(x){
  rowSums(as.matrix(x))
  })

n = nrow(AdjacencyList[[1]])
```

```{r, echo=FALSE}

hemisphere = vertex.attributes(graphList[[1]])$hemisphere


combinedDegreeList = lapply(DegreeList, function(x)x[hemisphere=="L"] + x[hemisphere=="R"])


```


```{r, echo=FALSE}
computeTV<-function(a,b){
  breaks = hist(c(a,b),breaks = 5,plot = F)$breaks
  ha = hist(a,breaks=breaks,plot = F)$density
  hb = hist(b,breaks=breaks,plot = F)$density
  ha = ha/sum(ha)
  hb = hb/sum(hb)
  sum(abs(ha - hb))/2
}

computeKS<- function(a,b){
  ks_test = ks.test(a,b,warning=FALSE)
  ks_test$statistic
}

plotDistance<- function(x,main="", q =0.95){
  plot(x,xlab="Vertex Pair Index",ylab="Distance",main=main)
  lines(x, col=((x>quantile(x, q)) +1), type="p")
}


getIndexOfLargeDistance<- function(x, q =0.95){
  c(1:length(x))[x>quantile(x, q)]
}

get3Distances<- function(a,b, n =n){
  tv<- sapply(c(1:n),function(i) computeTV(a[,i],b[,i]))
  ws<- sapply(c(1:n),function(i)  wasserstein1d(a[,i],b[,i]))
  ks<- sapply(c(1:n),function(i)  computeKS(a[,i],b[,i]))
  list("TV"=tv,"WS"=ws,"KS"=ks)
}

```

#Stastistical distance based vertex selection

In this report, we consider the matching ROIs in the left and right hemispheres as the smallest unit, and use distance based selection to choose the important pairs of ROIs that are associated with the traits of interests. We first calculate the sum of the degreess of two mathcing vertices (ROIs) in each pair, then compute the distance of the distribution of the degree sum between two groups that are compared.

The pair index (i) can be converted to vertix indices (j) by   j=(2*i-1, 2*i)


##Distance between degree distributions


### 1. Distances between SEX

Here are the plots of distances over vertice pairs, using three types of distance.

```{r}

degreeMat<- do.call("rbind",combinedDegreeList)
d1 =degreeMat[covariates$GENDER==1,]
d2 =degreeMat[covariates$GENDER==2,]

dist = get3Distances(d1,d2, n = n/2)

par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")


```

Here are the lists of important vertice pairs, selected by three different distances:

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

### 2. Distances between GENOTYPES


```{r}
degreeMat<- do.call("rbind",combinedDegreeList)
d1 =degreeMat[covariates$GENOTYPE==0,]
d2 =degreeMat[covariates$GENOTYPE==1,]
d3 =degreeMat[covariates$GENOTYPE==2,]

```


1. Genotype 0 vs 1
```{r}
dist = get3Distances(d1,d2,n/2)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertice pair

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

2. Genotype 0 vs 2
```{r}
dist = get3Distances(d1,d3,n/2)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertice pair

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

3. Genotype 1 vs 3
```{r}
dist = get3Distances(d2,d3,n/2)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertice pair

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

### 3. Distances between Age Groups

We first use k-means to obtain 3 age groups (with min and max age):


```{r}

# kmAge = kmeans(covariates$AGE_WEEKS,centers = 3)

# ageGroups0 = kmAge$cluster
# ageGroups = ageGroups0
# centerOrder = order(kmAge$centers)

ageGroups=  1 + (covariates$AGE_WEEKS>61)*1 + (covariates$AGE_WEEKS>88)*1


d1 =degreeMat[ageGroups==1,]
d2 =degreeMat[ageGroups==2,]
d3 =degreeMat[ageGroups==3,]


range(covariates$AGE_WEEKS[ageGroups==1])
range(covariates$AGE_WEEKS[ageGroups==2])
range(covariates$AGE_WEEKS[ageGroups==3])

```



1. Age group 1 vs 2
```{r}
dist = get3Distances(d1,d2,n/2)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertice pair

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

2. Age group 1 vs 3
```{r}
dist = get3Distances(d1,d3,n/2)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertice pair

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

3. Age group 2 vs 3
```{r}
dist = get3Distances(d2,d3,n/2)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertice pair

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```
