---
output:
  pdf_document: default
  html_document: default
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
require('igraph')
require('ggplot2')
require('reshape')
require('lsr')
require('transport')

setwd("c:/Users/leo/git/ABadea/")

listGs<- list.files(path = "./graphml/", pattern = "*.graphml")

#read in covariates and graph list
#find those with common ids, sort by id

covariates<- read.csv("./covariates/predictors.csv",stringsAsFactors = F)
ids <- unlist( lapply(listGs,function(x)strtrim(x,6)))
common_id<- intersect(covariates$RUNNO , ids)

covariates <- covariates[covariates$RUNNO%in%common_id,]
covariates <- covariates[order(covariates$RUNNO),]  

listGs<- listGs[ids%in%common_id]
listGs<- listGs[order(listGs)]

graphList<- lapply(listGs, function(x){
  read.graph( file = paste("./graphml/",x,sep = ""),format = "graphml")
})

AdjacencyList<- lapply(graphList, function(x){
  get.adjacency(x)
})

HemisphereList<- lapply(graphList, function(x){
  get.vertex.attribute(x,name="hemisphere")
})

DegreeList<- lapply(AdjacencyList, function(x){
  rowSums(as.matrix(x))
  })

n = nrow(AdjacencyList[[1]])
```

```{r, echo=FALSE}
computeTV<-function(a,b){
  breaks = hist(c(a,b),breaks = 5,plot = F)$breaks
  ha = hist(a,breaks=breaks,plot = F)$density
  hb = hist(b,breaks=breaks,plot = F)$density
  ha = ha/sum(ha)
  hb = hb/sum(hb)
  sum(abs(ha - hb))/2
}

computeKS<- function(a,b){
  ks_test = ks.test(a,b,warning=FALSE)
  ks_test$statistic
}

plotDistance<- function(x,main="", q =0.95){
  plot(x,xlab="Vertex Index",ylab="Distance",main=main)
  lines(x, col=((x>quantile(x, q)) +1), type="p")
}


getIndexOfLargeDistance<- function(x, q =0.95){
  c(1:length(x))[x>quantile(x, q)]
}

get3Distances<- function(a,b){
  tv<- sapply(c(1:n),function(i) computeTV(a[,i],b[,i]))
  ws<- sapply(c(1:n),function(i)  wasserstein1d(a[,i],b[,i]))
  ks<- sapply(c(1:n),function(i)  computeKS(a[,i],b[,i]))
  list("TV"=tv,"WS"=ws,"KS"=ks)
}

```



##Distance between degree distributions

To select the important vertices associated with the between-group difference, we first obtain the degree distribution for each group.

Over each vertex, we have two distributions for every pair of groups (e.g. SEX=1 vs SEX=2 over vertex 1). We can compute their between-group distance. Through this, we can get the distances over all vertices. The ones with the largest distances are likely the important vertices. In this case, we will pick the vertices with distances in the top 5%.

There are 3 types of popular nonparametric distances: Total variation, Wasserstein and Kolmogorov-Smirnov. Obviously, there is no best choice, but one sensible approach would be picking the distance by which the top 5% appear very different from the remaining 95%. We will illustrate and explain in the first analysis related to sex.



### 1. Distances between SEX

Here are the plots of distances over vertices, using three types of distance.

```{r}

degreeMat<- do.call("rbind",DegreeList)
d1 =degreeMat[covariates$GENDER==1,]
d2 =degreeMat[covariates$GENDER==2,]

dist = get3Distances(d1,d2)

par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")


```

Three distances produce three different sets of top vertices. But, in Wasserstein, the top few appear to be farther apart from the remaining, compared against the other two distances. That is, it gives a clearer separation of the top from the others. So Wasserstein would be the recommended choice in this case.

Here are the lists of important vertices, selected by three different distances:

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

### 2. Distances between GENOTYPES


```{r}
degreeMat<- do.call("rbind",DegreeList)
d1 =degreeMat[covariates$GENOTYPE==0,]
d2 =degreeMat[covariates$GENOTYPE==1,]
d3 =degreeMat[covariates$GENOTYPE==2,]

```


1. Genotype 0 vs 1
```{r}
dist = get3Distances(d1,d2)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertices

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

2. Genotype 0 vs 2
```{r}
dist = get3Distances(d1,d3)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertices

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

3. Genotype 1 vs 3
```{r}
dist = get3Distances(d2,d3)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertices

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

### 3. Distances between Age Groups

We first use k-means to obtain 3 age groups (with min and max age):


```{r}

kmAge = kmeans(covariates$AGE_WEEKS,centers = 3)

ageGroups0 = kmAge$cluster
ageGroups = ageGroups0
centerOrder = order(kmAge$centers)

for(i in 1:3){
  ageGroups[ ageGroups0 ==centerOrder[i]] = i
}


d1 =degreeMat[ageGroups==1,]
d2 =degreeMat[ageGroups==2,]
d3 =degreeMat[ageGroups==3,]


range(covariates$AGE_WEEKS[ageGroups==1])
range(covariates$AGE_WEEKS[ageGroups==2])
range(covariates$AGE_WEEKS[ageGroups==3])

```



1. Age group 1 vs 2
```{r}
dist = get3Distances(d1,d2)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertices

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

2. Age group 1 vs 3
```{r}
dist = get3Distances(d1,d3)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertices

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```

3. Age group 2 vs 3
```{r}
dist = get3Distances(d2,d3)
par(mfrow=c(1,3))
plotDistance(dist$TV, "Total Variation")
plotDistance(dist$WS, "Wasserstein")
plotDistance(dist$KS, "Kolmogorov-Smirnov")
```

Important vertices

Total variation
```{r}
getIndexOfLargeDistance(dist$TV)
```
Wasserstein
```{r}
getIndexOfLargeDistance(dist$WS)
```
Kolmogorov-Smirnov
```{r}
getIndexOfLargeDistance(dist$KS)
```
